Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
adjust pics in the power cmp

1. experiments -> data collection periods

@@

In the following analysis, lineups with uniform fitted values will be the focus. Visual patterns are more likely to be revealed under a uniform distribution. Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis. Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-other-fitted-value-distributions}.

@@

For non-linearity patterns, the power curve of RESET test climbs aggressively from
`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered. Meanwhile, no noticeable visual features can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is also almost always greater than the power of visual test. For $0 \leq log_e(E) \leq 2$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$. Similarly, the visual feature is nearly unobservable from the example residual plots.

@@

set.seed(106)

@@

# Obtain data at different effect. Reuse x and recompute e.
for (i in target_sigma) {
  j <- j + 1
plist[[j]] <- poly_model(shape = 2, 
                         x = rand_uniform())$
                  gen(n = 100, 
                      computed =  ex_dat %>%
                        mutate(e = sqrt(i) * e) %>%
                        select(x, e)
                      ) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5), 
                remove_axis = TRUE, 
                remove_grid_line = TRUE, 
                size = 0.05) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

@@

set.seed(10091)

@@

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 1, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.05) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

>>>>

@@

In the following analysis, lineups with uniform fitted values will be the focus. Visual patterns are more likely to be revealed under a uniform distribution. Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis. Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.

@@

For non-linearity patterns, the power curve of RESET test climbs aggressively from
`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered. Meanwhile, no noticeable residual depatures can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is also almost always greater than the power of visual test. For $log_e(E)$ at around 2.5, where the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$. Similarly, the visual feature is nearly unobservable from the example residual plots.

@@

# 106
# 112
# 113
# 115
# 132
# 133
# 144!
# 147
set.seed(132)

@@

# Obtain data at different effect. Reuse x and recompute e.
for (i in target_sigma) {
  j <- j + 1
plist[[j]] <- poly_model(shape = 2, 
                         x = rand_uniform())$
                  gen(n = 100, 
                      computed = ex_dat %>%
                        mutate(e = sqrt(i) * e) %>%
                        select(x, e)
                      ) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5), 
                remove_axis = TRUE, 
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

@@

# 10091
# 101
# 105
# 106
# 112
# 119
# 133
set.seed(135)

@@

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 1, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
adjust pics in the power cmp

Instead, we directly estimate the probability of a lineup being rejected
by assuming that individual skill has negligible effect on the variation
of the power. The assumption is to simplify the model structure, thereby
obviate costly large-scale experiments to estimate complex covariance
matrices. The model is a logistic regression with the natural logarithm
of the effect as the only regressor formulated as

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log(E)),
\end{equation}

\noindent where \(\Lambda(.)\) is the standard logistic function given
as \(\Lambda(z) = exp(z)/(1+exp(z))\).

Effect \(E\) is derived from the Kullback-Leibler divergence (see
Appendix for the details) formulated as

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  experiments -\textgreater{} data collection periods
\end{enumerate}

@@

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 7\% to 30\% as \(log_e(E)\) increases from 0 to 2,
while power of other tests respond inactively to the change of effect,
showing that RESET test is way more sensitive to the type of model
defects that being considered. Meanwhile, no noticeable visual features
can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
is also almost always greater than the power of visual test. For
\(0 \leq log_e(E) \leq 2\), where the power curve of the visual test
remains at a low level, the Breusch-Pagan test still have a decent
amount of chance of rejecting \(H_0\). Similarly, the visual feature is
nearly unobservable from the example residual plots.

>>>>

Instead, we directly estimate the probability of a lineup being rejected
by assuming that individual skill has negligible effect on the variation
of the power. The assumption is to simplify the model structure, thereby
obviate costly large-scale experiments to estimate complex covariance
matrices. The model is a logistic regression formulated as

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(log(\frac{0.5}{0.95}) + \beta_1 E),
\end{equation}

\noindent where \(\Lambda(.)\) is the standard logistic function given
as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size is the only
regressor and the intercept is fixed to \(log(0.5/0.95)\) so that
\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).

Effect size \(E\) is derived from the Kullback-Leibler divergence (see
Appendix for the details) formulated as

@@

@@

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 7\% to 63\% as \(log_e(E)\) increases from 0 to 2.5,
while power of other tests respond inactively to the change of effect,
showing that RESET test is way more sensitive to the type of model
defects that being considered. Meanwhile, no noticeable residual
depatures can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
is also almost always greater than the power of visual test. For
\(log_e(E)\) at around 2.5, where the power curve of the visual test
remains at a low level, the Breusch-Pagan test still have a decent
amount of chance of rejecting \(H_0\). Similarly, the visual feature is
nearly unobservable from the example residual plots.
